---
title: "Discharge Data"
author: "Sasha Keller"
date: "4/12/2022"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
# Set your working directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# Set Output options
knitr::opts_chunk$set(echo=FALSE)

# Load your packages
library(EcoHydRology)
library(tidyverse)
library(cowplot)
library(dataRetrieval)
library(lubridate)
library(lfstat)
library(trend)
library(gt)
library(scales)
# Set your ggplot theme
theme_set(theme_classic())
# Load your datasets

```
```{r, process ENSO data}
enso<-read_csv("Data/Raw/monthly_enso.csv")
enso$Months<-as.Date(enso$Months, form="%y/%m/%d")
enso_processed<-enso%>%
  mutate(Year = year(Months),
         Month = month(Months))%>%
  group_by(Year, Month) %>%
  mutate(season.class=case_when(SST>0.5 ~ "El Niño",
                                SST< -0.5 ~ "La Niña",
                                TRUE~"Neutral"))
```

```{r} 
#Tulalip
TulalipDischarge <- readNWISdv(siteNumbers = "12158040",
                     parameterCd = "00060", # discharge (ft3/s)
                     startDate = "1974-10-01", 
                     endDate = "2021-09-30")
names(TulalipDischarge)[4:5] <- c("Discharge", "Approval.Code")
attr(TulalipDischarge, "variableInfo")
attr(TulalipDischarge, "siteInfo")

TulalipProccessed <- TulalipDischarge %>%
  mutate(Baseflow = baseflow(Discharge), 
         Stormflow = Discharge - Baseflow,
         Year = year(Date), 
         Year = year(Date),
         Month = month(Date))
TulalipENSO<-TulalipProccessed%>%
    filter(Year>1999)%>%
  left_join(enso_processed)

TulalipMonthly<- TulalipProccessed%>%
  group_by(Year, Month) %>%
  summarise(Discharge_acftmo = sum(Discharge)*1.98347)

TulalipENSO<-TulalipProccessed%>%
    filter(Year>1999)%>%
  left_join(enso_processed)%>%
  left_join(TulalipMonthly)

#Puyallup
PuyallupDischarge <- readNWISdv(siteNumbers = "12061500",
                     parameterCd = "00060", # discharge (ft3/s)
                     startDate = "1974-10-01", 
                     endDate = "2021-09-30")

names(PuyallupDischarge)[4:5] <- c("Discharge", "Approval.Code")
attr(PuyallupDischarge, "variableInfo")
attr(PuyallupDischarge, "siteInfo")

PuyallupProccessed <- PuyallupDischarge %>%
  mutate(Baseflow = baseflow(Discharge), 
         Stormflow = Discharge - Baseflow,
         Year = year(Date), 
         Year = year(Date),
         Month = month(Date))

PuyallupMonthly<- PuyallupProccessed%>%
  group_by(Year, Month) %>%
  summarise(Discharge_acftmo = sum(Discharge)*1.98347)

PuyallupENSO<-PuyallupProccessed%>%
    filter(Year>1999)%>%
  left_join(enso_processed)%>%
  left_join(PuyallupMonthly)

#Duckabush
DuckabushDischarge <- readNWISdv(siteNumbers = "12054000",
                     parameterCd = "00060", # discharge (ft3/s)
                     startDate = "1974-10-01", 
                     endDate = "2021-09-30")
names(DuckabushDischarge)[4:5] <- c("Discharge", "Approval.Code")
attr(DuckabushDischarge, "variableInfo")
attr(DuckabushDischarge, "siteInfo")

DuckProccessed <- DuckabushDischarge %>%
  mutate(Baseflow = baseflow(Discharge), 
         Stormflow = Discharge - Baseflow,
         Year = year(Date), 
         Year = year(Date),
         Month = month(Date))

DuckMonthly<- DuckProccessed%>%
  group_by(Year, Month) %>%
  summarise(Discharge_acftmo = sum(Discharge)*1.98347)

DuckabushENSO<-DuckProccessed%>%
    filter(Year>1999)%>%
  left_join(enso_processed)%>%
  left_join(DuckMonthly)

#Skokomish
SkokomishDischarge <- readNWISdv(siteNumbers = "12101500",
                     parameterCd = "00060", # discharge (ft3/s)
                     startDate = "1974-10-01", 
                     endDate = "2021-09-30")

names(SkokomishDischarge)[4:5] <- c("Discharge", "Approval.Code")
attr(SkokomishDischarge, "variableInfo")
attr(SkokomishDischarge, "siteInfo")

SkokomishProccessed <- SkokomishDischarge %>%
  mutate(Baseflow = baseflow(Discharge), 
         Stormflow = Discharge - Baseflow,
         Year = year(Date), 
         Year = year(Date),
         Month = month(Date))
SkokomishENSO<-SkokomishProccessed%>%
    filter(Year>1999)%>%
  left_join(enso_processed)

SkokomishMonthly<- SkokomishProccessed%>%
  group_by(Year, Month) %>%
  summarise(Discharge_acftmo = sum(Discharge)*1.98347)

SkokomishENSO<-SkokomishProccessed%>%
    filter(Year>1999)%>%
  left_join(enso_processed)%>%
  left_join(SkokomishMonthly)

  
ggplot(PuyallupENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point() +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Puyallup Daily Discharge")

ggplot(TulalipENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point() +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Tulalip Daily Discharge")

ggplot(SkokomishENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point() +
  geom_smooth()+
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Skokomish Daily Discharge")


```



```{r, Duckabush}
#Climate events will be identified using NOAA's Climate Prediction Center (https://origin.cpc.ncep.noaa.gov/products/analysis_monitoring/ensostuff/ONI_v5.php).
#El Nino events should create a warmer, drier climate in Washington while La Nina will create a cooler, wetter environment. 


DuckLa<- DuckabushENSO %>%
  filter(season.class%in% c("La Niña",  "Neutral"))
DuckEl<-DuckabushENSO %>%
  filter(season.class%in% c("Neutral",  "El Niño"))
DuckEx<-DuckabushENSO %>%
  filter(season.class %in% c("La Niña",  "El Niño"))


#Is there  a significant difference in monthly discharges between season classes?
DuckKTMonthly<-kruskal.test(Discharge_acftmo ~ season.class, data = DuckabushENSO)
#Is stormflow different between Neutral and El Nino?
DuckWTMEl<-wilcox.test(Discharge_acftmo ~ season.class, data=DuckEl)
#Is stormflow different between Neutral and La Nina?
DuckWTMLa<-wilcox.test(Discharge_acftmo ~ season.class, data=DuckLa)
#Is stormflow different between La Nina and El Nino?
DuckWTMEx<-wilcox.test(Discharge_acftmo ~ season.class, data=DuckEx)

ggplot(DuckabushENSO, aes(x=Year,y=Discharge_acftmo, color=season.class))+
  geom_point(size=4) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  scale_y_log10()+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Duckabush Monthly Discharge")



#How about daily discharge?
DuckKTDaily<-kruskal.test(Discharge ~ season.class, data = DuckabushENSO)
#Is stormflow different between Neutral and El Nino?
DuckWTDEl<-wilcox.test(Discharge ~ season.class, data=DuckEl)
#Is stormflow different between Neutral and La Nina?
DuckWTDLa<-wilcox.test(Discharge ~ season.class, data=DuckLa)
#Is stormflow different between La Nina and El Nino?
DuckWTDEx<-wilcox.test(Discharge ~ season.class, data=DuckEx)

ggplot(DuckabushENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Duckabush Daily Discharge")

 #What about during the largest precipitation events?
DuckStormKT<-kruskal.test(Stormflow ~ season.class, data=DuckabushENSO)
#Is stormflow different between Neutral and El Nino?
DuckStormWEl<-wilcox.test(Stormflow ~ season.class, data=DuckEl)
#Is stormflow different between Neutral and La Nina?
DuckStormWLa<-wilcox.test(Stormflow ~ season.class, data=DuckLa)
#Is stormflow different between La Nina and El Nino?
DuckStormWEx<-wilcox.test(Stormflow ~ season.class, data=DuckEx)

DuckMonthlyStats<-data.frame(
  DuckWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(DuckWTMEl$p.value, DuckWTMLa$p.value, DuckWTMEx$p.value))

DuckDailyStats<-data.frame(
  DuckWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(DuckWTDEl$p.value, DuckWTDLa$p.value, DuckWTDEx$p.value))

DuckStormStats<-data.frame(
  DuckWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(DuckStormWEl$p.value, DuckStormWLa$p.value, DuckStormWEx$p.value))


```

```{r, Puyallup}

PuyallupLa<- PuyallupENSO %>%
  filter(season.class%in% c("La Niña",  "Neutral"))
PuyallupEl<-PuyallupENSO %>%
  filter(season.class%in% c("Neutral",  "El Niño"))
PuyallupEx<-PuyallupENSO %>%
  filter(season.class %in% c("La Niña",  "El Niño"))


#Is there  a significant difference in monthly discharges between season classes?
PuyallupKTMonthly<-kruskal.test(Discharge_acftmo ~ season.class, data = PuyallupENSO)
#Is stormflow different between Neutral and El Nino?
PuyallupWTMEl<-wilcox.test(Discharge_acftmo ~ season.class, data=PuyallupEl)
#Is stormflow different between Neutral and La Nina?
PuyallupWTMLa<-wilcox.test(Discharge_acftmo ~ season.class, data=PuyallupLa)
#Is stormflow different between La Nina and El Nino?
PuyallupWTMEx<-wilcox.test(Discharge_acftmo ~ season.class, data=PuyallupEx)

ggplot(PuyallupENSO, aes(x=Year,y=Discharge_acftmo, color=season.class))+
  geom_point(size=4) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Puyallup Monthly Discharge")



#How about daily discharge?
PuyallupKTDaily<-kruskal.test(Discharge ~ season.class, data = PuyallupENSO)
#Is stormflow different between Neutral and El Nino?
PuyallupWTDEl<-wilcox.test(Discharge ~ season.class, data=PuyallupEl)
#Is stormflow different between Neutral and La Nina?
PuyallupWTDLa<-wilcox.test(Discharge ~ season.class, data=PuyallupLa)
#Is stormflow different between La Nina and El Nino?
PuyallupWTDEx<-wilcox.test(Discharge ~ season.class, data=PuyallupEx)

ggplot(PuyallupENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Puyallup Daily Discharge")

 #What about during the largest precipitation events?
PuyallupStormKT<-kruskal.test(Stormflow ~ season.class, data=PuyallupENSO)
#Is stormflow different between Neutral and El Nino?
PuyallupStormWEl<-wilcox.test(Stormflow ~ season.class, data=PuyallupEl)
#Is stormflow different between Neutral and La Nina?
PuyallupStormWLa<-wilcox.test(Stormflow ~ season.class, data=PuyallupLa)
#Is stormflow different between La Nina and El Nino?
PuyallupStormWEx<-wilcox.test(Stormflow ~ season.class, data=PuyallupEx)

PuyallupMonthlyStats<-data.frame(
  PuyallupWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(PuyallupWTMEl$p.value, PuyallupWTMLa$p.value, PuyallupWTMEx$p.value))

PuyallupDailyStats<-data.frame(
  PuyallupWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(PuyallupWTDEl$p.value, PuyallupWTDLa$p.value, PuyallupWTDEx$p.value))

PuyallupStormStats<-data.frame(
  PuyallupWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(PuyallupStormWEl$p.value, PuyallupStormWLa$p.value, PuyallupStormWEx$p.value))

```

```{r, Skokomish}
SkokomishLa<- SkokomishENSO %>%
  filter(season.class%in% c("La Niña",  "Neutral"))
SkokomishEl<-SkokomishENSO %>%
  filter(season.class%in% c("Neutral",  "El Niño"))
SkokomishEx<-SkokomishENSO %>%
  filter(season.class %in% c("La Niña",  "El Niño"))


#Is there  a significant difference in monthly discharges between season classes?
SkokomishKTMonthly<-kruskal.test(Discharge_acftmo ~ season.class, data = SkokomishENSO)
#Is stormflow different between Neutral and El Nino?
SkokomishWTMEl<-wilcox.test(Discharge_acftmo ~ season.class, data=SkokomishEl)
#Is stormflow different between Neutral and La Nina?
SkokomishWTMLa<-wilcox.test(Discharge_acftmo ~ season.class, data=SkokomishLa)
#Is stormflow different between La Nina and El Nino?
SkokomishWTMEx<-wilcox.test(Discharge_acftmo ~ season.class, data=SkokomishEx)

ggplot(SkokomishENSO, aes(x=Year,y=Discharge_acftmo, color=season.class))+
  geom_point(size=4) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  scale_y_log10()+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Skokomish Monthly Discharge")



#How about daily discharge?
SkokomishKTDaily<-kruskal.test(Discharge ~ season.class, data = SkokomishENSO) 
#Is stormflow different between Neutral and El Nino?
SkokomishWTDEl<-wilcox.test(Discharge ~ season.class, data=SkokomishEl)
#Is stormflow different between Neutral and La Nina?
SkokomishWTDLa<-wilcox.test(Discharge ~ season.class, data=SkokomishLa)
#Is stormflow different between La Nina and El Nino?
SkokomishWTDEx<-wilcox.test(Discharge ~ season.class, data=SkokomishEx)

ggplot(SkokomishENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)")+
  ggtitle("Skokomish Daily Discharge")

 #What about during the largest precipitation events?
SkokomishStormKT<-kruskal.test(Stormflow ~ season.class, data=SkokomishENSO)
#Is stormflow different between Neutral and El Nino?
SkokomishStormWTEl<-wilcox.test(Stormflow ~ season.class, data=SkokomishEl)
#Is stormflow different between Neutral and La Nina?
SkokomishStormWTLa<-wilcox.test(Stormflow ~ season.class, data=SkokomishLa)
#Is stormflow different between La Nina and El Nino?
SkokomishStormWTEx<-wilcox.test(Stormflow ~ season.class, data=SkokomishEx)

SkokomishMonthlyStats<-data.frame(
  SkokomishWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(SkokomishWTMEl$p.value, SkokomishWTMLa$p.value, SkokomishWTMEx$p.value))

SkokomishDailyStats<-data.frame(
  SkokomishWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(SkokomishWTDEl$p.value, SkokomishWTDLa$p.value, SkokomishWTDEx$p.value))

SkokomishStormStats<-data.frame(
  SkokomishWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(SkokomishStormWTEl$p.value, SkokomishStormWTLa$p.value, SkokomishStormWTEx$p.value))

```

```{r, Tulalip}
TulalipLa<- TulalipENSO %>%
  filter(season.class%in% c("La Niña",  "Neutral"))
TulalipEl<-TulalipENSO %>%
  filter(season.class%in% c("Neutral",  "El Niño"))
TulalipEx<-TulalipENSO %>%
  filter(season.class %in% c("La Niña",  "El Niño"))


#Is there  a significant difference in monthly discharges between season classes?
TulalipKTMonthly<-kruskal.test(Discharge_acftmo ~ season.class, data = TulalipENSO)
#Is stormflow different between Neutral and El Nino?
TulalipWTMEl<-wilcox.test(Discharge_acftmo ~ season.class, data=TulalipEl)
#Is stormflow different between Neutral and La Nina?
TulalipWTMLa<-wilcox.test(Discharge_acftmo ~ season.class, data=TulalipLa)
#Is stormflow different between La Nina and El Nino?
TulalipWTMEx<-wilcox.test(Discharge_acftmo ~ season.class, data=TulalipEx)




#How about daily discharge?
TulalipKTDaily<-kruskal.test(Discharge ~ season.class, data = TulalipENSO)
#Is stormflow different between Neutral and El Nino?
TulalipWTDEl<-wilcox.test(Discharge ~ season.class, data=TulalipEl)
#Is stormflow different between Neutral and La Nina?
TulalipWTDLa<-wilcox.test(Discharge ~ season.class, data=TulalipLa)
#Is stormflow different between La Nina and El Nino?
TulalipWTDEx<-wilcox.test(Discharge ~ season.class, data=TulalipEx)

 #What about during the largest precipitation events?
TulalipStormKT<-kruskal.test(Stormflow ~ season.class, data=TulalipENSO)
#Is stormflow different between Neutral and El Nino?
TulalipStormWTEl<-wilcox.test(Stormflow ~ season.class, data=TulalipEl)
#Is stormflow different between Neutral and La Nina?
TulalipStormWTLa<-wilcox.test(Stormflow ~ season.class, data=TulalipLa)
#Is stormflow different between La Nina and El Nino?
TulalipStormWTEx<-wilcox.test(Stormflow ~ season.class, data=TulalipEx)

TulalipMonthlyStats<-data.frame(
  TulalipWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(TulalipWTMEl$p.value, TulalipWTMLa$p.value, TulalipWTMEx$p.value))

TulalipDailyStats<-data.frame(
  TulalipWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(TulalipWTDEl$p.value, TulalipWTDLa$p.value, TulalipWTDEx$p.value))

TulalipStormStats<-data.frame(
  TulalipWilcoxonTests = c("El Nino ~ Neutral", "La Nina ~ Neutral", "El Nino ~ La Nina"),
  pValue= c(TulalipStormWTEl$p.value, TulalipStormWTLa$p.value, TulalipStormWTEx$p.value))
```
```{r}

KTMonthly<-data.frame(
  StreamName=c("Duckabush","Skokomish","Puyallup","Tulalip"),
               pValue=c(DuckKTMonthly$p.value, SkokomishKTMonthly$p.value,PuyallupKTMonthly$p.value,TulalipKTMonthly$p.value))

KTDaily<-data.frame(
  StreamName=c("Duckabush","Skokomish","Puyallup","Tulalip"),
               pValue=c(DuckKTDaily$p.value, SkokomishKTDaily$p.value,PuyallupKTDaily$p.value,TulalipKTDaily$p.value))

KTStorm<-data.frame(
  StreamName=c("Duckabush","Skokomish","Puyallup","Tulalip"),
               pValue=c(DuckStormKT$p.value, SkokomishStormKT$p.value,PuyallupStormKT$p.value,TulalipStormKT$p.value))


KTMonthlytable<-KTMonthly%>%
  gt()%>%
  tab_header(title = md("**Monthly Discharge Kruskal-Wallis Rank Sum Results**"))%>%
  cols_label(
    StreamName = md("**Stream**"),
    pValue = md("**p Value**"))
KTMonthlytable

KTDailytable<-KTDaily%>%
  gt()%>%
  tab_header(title = md("**Daily Discharge Kruskal-Wallis Rank Sum Results**"))%>%
  cols_label(
    StreamName = md("**Stream**"),
    pValue = md("**p Value**"))
KTDailytable

KTStormtable<-KTStorm%>%
  gt()%>%
  tab_header(title = md("**Storm Discharge Kruskal-Wallis Rank Sum Results**"))%>%
  cols_label(
    StreamName = md("**Stream**"),
    pValue = md("**p Value**"))
KTStormtable
```


```{r, Puyallup}
#Base tables
table1 = PuyallupMonthlyStats

table2 = PuyallupDailyStats

table3=PuyallupStormStats

#If the individual tables are a GT object already, exact the data
#gtTable = gt(table1)
#table1 = gtTable$`_data`


#Merge tables by row
fullTable = rbind(table1, table2,table3)
fullTable

PuyallupgtTable = gt(fullTable) %>%
  tab_header(title = md("**Puyallup Wilcoxon Rank Sum Results**"))%>%
  cols_label(
    PuyallupWilcoxonTests =  md("**Seasons Compared**"),
    pValue =  md("**p Value**"))%>%
  tab_row_group(
    group = "Monthly Discharge",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Daily Discharge",
    rows = 4:6
  )%>% 
  tab_row_group(
    group = "Storm Event Discharge",
    rows = 7:9
  )

PuyallupgtTable
```
```{r, Duck}
#Base tables
table1 = DuckMonthlyStats

table2 = DuckDailyStats

table3=DuckStormStats

#If the individual tables are a GT object already, exact the data
#gtTable = gt(table1)
#table1 = gtTable$`_data`


#Merge tables by row
fullTable = rbind(table1, table2,table3)
fullTable

DuckgtTable = gt(fullTable) %>%
  tab_header(title = md("**Duckabush Wilcoxon Rank Sum Results**"))%>%
  cols_label(
    DuckWilcoxonTests = md("**Seasons Compared**"),
    pValue = md("**p Value**"))%>%
  tab_row_group(
    group = "Monthly Discharge",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Daily Discharge",
    rows = 4:6
  )%>% 
  tab_row_group(
    group = "Storm Event Discharge",
    rows = 7:9
  )

DuckgtTable
```
```{r, Tulalip}
#Base tables
table1 = TulalipMonthlyStats

table2 = TulalipDailyStats

table3=TulalipStormStats

#If the individual tables are a GT object already, exact the data
#gtTable = gt(table1)
#table1 = gtTable$`_data`


#Merge tables by row
fullTable = rbind(table1, table2,table3)
fullTable

TulalipgtTable = gt(fullTable) %>%
  tab_header(title = md("**Tulalip Wilcoxon Rank Sum Results**"))%>%
  cols_label(
    TulalipWilcoxonTests = md("**Seasons Compared**"),
    pValue = md("**p Value**"))%>%
  tab_row_group(
    group = "Monthly Discharge",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Daily Discharge",
    rows = 4:6
  )%>% 
  tab_row_group(
    group = "Storm Event Discharge",
    rows = 7:9
  )

TulalipgtTable
```
```{r,Skokomish}
#Base tables
table1 = SkokomishMonthlyStats

table2 = SkokomishDailyStats

table3=SkokomishStormStats

#If the individual tables are a GT object already, exact the data
#gtTable = gt(table1)
#table1 = gtTable$`_data`


#Merge tables by row
fullTable = rbind(table1, table2,table3)
fullTable

SkokomishgtTable = gt(fullTable) %>%
  tab_header(title = md("**Skokomish Wilcoxon Rank Sum Results**"))%>%
  cols_label(
    SkokomishWilcoxonTests = md("**Seasons Compared**"),
    pValue = md("**p Value**"))%>%
  tab_row_group(
    group = "Monthly Discharge",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Daily Discharge",
    rows = 4:6
  )%>% 
  tab_row_group(
    group = "Storm Event Discharge",
    rows = 7:9
  )

SkokomishgtTable
```
```{r,export data?}
saveRDS(SkokomishgtTable, PuyallupgtTable, DuckgtTable,TulalipgtTable, file="~/Data/Processed/figures.RData")
#KTMonthlytable,KTDailytable,KTStormtable,
```

```{r, discharge plots}
#Duckabush
ggplot(DuckabushENSO, aes(x=Date,y=Discharge_acftmo, color=season.class))+
  geom_point(size=3) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Duckabush Monthly Discharge")+
  scale_y_continuous(labels= comma)


ggplot(DuckabushENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Duckabush Daily Discharge")+
  scale_y_continuous(labels= comma)

ggplot(DuckabushENSO, aes(x = Date, y = Stormflow, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Duckabush Stormflow Discharge")+
  scale_y_continuous(labels= comma)

#Puyallup
ggplot(PuyallupENSO, aes(x=Year,y=Discharge_acftmo, color=season.class))+
  geom_point(size=4) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  scale_y_log10()+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Puyallup Monthly Discharge")+
  scale_y_continuous(labels= comma)


ggplot(PuyallupENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Puyallup Daily Discharge")+
  scale_y_continuous(labels= comma)

ggplot(PuyallupENSO, aes(x = Date, y = Stormflow, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Puyallup Stormflow Discharge")+
  scale_y_continuous(labels= comma)

#Tulalip
ggplot(TulalipENSO, aes(x=Year,y=Discharge_acftmo, color=season.class))+
  geom_point(size=4) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  scale_y_log10()+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Tulalip Monthly Discharge")+
  scale_y_continuous(labels= comma)


ggplot(TulalipENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Tulalip Daily Discharge")+
  scale_y_continuous(labels= comma)

ggplot(TulalipENSO, aes(x = Date, y = Stormflow, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Tulalip Stormflow Discharge")+
  scale_y_continuous(labels= comma)

#Skokomish
ggplot(SkokomishENSO, aes(x=Year,y=Discharge_acftmo, color=season.class))+
  geom_point(size=4) +
  
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Skokomish Monthly Discharge")+
  scale_y_continuous(labels= comma)


ggplot(SkokomishENSO, aes(x = Date, y = Discharge, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Skokomish Daily Discharge")+
  scale_y_continuous(labels= comma)

ggplot(SkokomishENSO, aes(x = Date, y = Stormflow, color=season.class)) +
  geom_point(alpha=0.7) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Skokomish Stormflow Discharge")+
  scale_y_continuous(labels= comma)

#Example

DuckabushSample<- DuckabushENSO %>%
  filter(Year > 2004 & Year < 2011)

ggplot(DuckabushSample, aes(x=Date,y=Discharge, color=season.class))+
  geom_point(size=4) +
    scale_color_manual(values = c("El Niño"="#feb24c",
                                  "La Niña"="#2c7fb8",
                                  "Neutral"="#7fcdbb"))+
  labs(x = "Year", y = "Discharge (cfs)",color="ENSO Season")+
  ggtitle("Duckabush Daily Discharge 2005-2010")+
  scale_y_continuous(labels= comma)
```
Questions: 
Is 3 month seasonal mean common water format or just climate data?


Causes for changes in discharge?
el nino la nina working as intended
el nino la nina 



How to identify relevant baseflow with so many ~*events*~ happening?




What is proper to compare?
code la nina, el nino, neither times to the data set then do summary stats to investigate? t test?


SMK t test for artifical seasons? 
Can I just recode El Nino/La Nina/Neither as 1,2,3 and run a t test frequency = 3?




#TO DO
- Put stat tests into variables, put variables into table
- Make a smaller time frame discharge plot to show variance in season?
- Put all ggplots together somehow
- save datasets as objects, so I can put the plot code in another file
- Finish write up






